---
title: "Train Collision"
format: html
editor: visual
---

# Train Collision: Displacement vs. Time Plot

## Background

-   Two trains are on the same track and separated by 100 miles.
-   They are heading towards each other at a speed of 50 mph.
-   When the trains depart, a bird flies off the first train and flies towards the 2nd train at 75 mph.
-   When the bird reaches the 2nd train, it turns around and continues back towards the first train at the same speed of 75 mph.
-   After reaching the 1st train again, it turns around and continues to the 2nd train. It continues going back and forth between the two trains until the trains collide.
-   How far does the bird travel before the two trains collide?

::: callout-note
The **peregrine falcon** is the fastest bird in the world. It can reach speeds of up to 240 mph in a high-speed dive, called a stoop. So a bird traveling 75 mph is entirely feasible.
:::

## The Plot

Displacement (miles) vs. Time (minutes)

-   x-axis range: (0 - 65?) (max time of graph at 60)
-   y-axis range (0 - 110?) (max range of graph is 100)

## Data Set:

> file: train_data_displacement_vs_time.csv \* One Graph with Two plots on same xy-axis

### Data Column Headers

-   x-axis: time_min
-   y-axis: 1st plot: train_1_displacement_miles
-   y-axis: 2nd plot: train_2_displacement_miles

## Rough Animation Flow

- [-]  Display Graph with starting point (as a dot) for each plot at time = 0 (first point). Label plot at origin as train 1. Label plot beginning at (0,100) as train 2.
- [-]  Slowly plot each simultaneously.
- [ ]  When plots intersect at the last point, indicate an explosion (cartoon drawn explosions?)
- [ ]  Animated dashed horizontal line starting from point of intersection (60,50) to displacement-axis (0,50)
- [ ]  Centered text (Segoe UI : Light) appears over dashed-line indicating 50 miles.
- [ ]  Animated dashed vertical line starting from point of intersection (60,50) to time-axis (0,50).
- [ ]  Text (Segoe UI : Light) appears to the left of the dashed-line indicating 1 hour.

## Coding

```{r setup}
# load libraries ----
library(tidyverse)
library(janitor)
library(extrafont)
library(gganimate)
library(gifski)
library(Cairo)
library(ggrepel)

# # font_import() # run this once
# loadfonts(device = "win")
CairoWin()

# load data ----
df <- read_csv("data/train_data_displacement_vs_time.csv")

theme_set(
  theme_minimal(base_size = 32) + 
    theme(
      text = element_text(family = "Segoe UI Light"), #, face = "italic")
      panel.border = element_rect(color = "black", linewidth = 1, fill = NA)
    )
)

```

```{r tidy}
df1 <- df |> 
  pivot_longer(cols = starts_with("train"),
               names_to = "train",
               names_pattern = "(.*)_displacement_miles",
               values_to = "displacement_miles") |> 
  mutate(train = train |> 
           str_replace("_", " ") |> 
           str_to_title())

df1 |> 
  ggplot(
    aes(x = time_min,
        y = displacement_miles,
        group = train,
        color = train,
        label = train)
  ) +
  geom_line(linewidth = 2, lineend = "round") +
  geom_point(size = 5) +
  guides(color = "none") +
  labs(
    x = "Time (min)",
    y = "Displacement (miles)"
  ) +
  geom_text_repel(hjust = 0, nudge_x = 3, size = 8) +
  coord_cartesian(clip = "off") +
  transition_reveal(time_min) +
  view_follow(fixed_x = TRUE)
```
